---
title: "P8106_hw2_xh2395"
author: "Xin  He"
date: "3/19/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
library(tidyverse)
library(caret)
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5) 
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1) 
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
library(patchwork)
library(splines)
library(gam)
library(mgcv)
library(boot)
library(ggplot2)
library(pdp)
library(earth)
```

## Homework 2 Description

In this exercise, we build nonlinear models using the "College" data. The dataset
contains statistics for 565 US Colleges from the 1995 issue of US News and World Report.
The response variable is the out-of-state tuition (Outstate). In what follows, use the data excluding statistics for Columbia University (i.e., the 125th observation) to train the models.

## Import the data

```{r}
college_df = read_csv("./data/College.csv")
```

## Set random seed

```{r}
set.seed(2020)
```


## Answer of the questions
